Optimal Approximation of Discrete Random Variables \\ for Estimation of Probabilities of Missing Deadlines

### 2018-09-05 Kolmogorov_Approx.tex
Updates from ShareLaTeX

@@

\DeclareMathOperator{\supp}{support}
\DeclareMathOperator{\support}{support}
\DeclareMathOperator{\Trim}{Trim}
\DeclareMathOperator{\LTrim}{LTrim}
\SetKwFunction{mEqOne}{mEqOne} 
\DeclareMathOperator{\KlmApprox}{Approx}
%\DeclareMathOperator{\OptTrim}{KolmogorovApprox}
\DeclareMathOperator{\OptTrim}{OptTrim}

  \pdfinfo{
	/Title (Deadline estimation by using an optimal approximation of discrete random variables with respect to the Kolmogorov distance
	)
	/Author ()}
\setcounter{secnumdepth}{0}

@@

Various approaches for approximation of probability distributions are studied in the literature~\cite{AMCR83,pavlikov2016cvar,PS77,vidyasagar2012metric,cohen2015estimating,CohenGW18}. 
These approaches vary in the types random variables considered, how they are represented, and in the criteria used for evaluation of the quality of the approximations. In this paper we propose an approach for compressing the probability mass function of a random variable $X$ such that the errors added to queries such as $Pr(X\leq t)$, for  any $t>0$, is minimal. In other words, we minimise the Kolmogorov distance between the approximation and the original random variable. Our motivation for this is estimation of the probability for missing deadlines, as described, e.g.,in~\cite{cohen2015estimating} and in~\cite{Kashef18}. Specifically, when $X$ represents the probability distribution of the time to complete some complex schedule and we cannot afford to maintain the full table of its probability mass function, we propose an algorithm for producing a smaller table, whose size can be specified, such that probabilities for missing deadlines are preserved as much as possible.

@@

The case study examined in this paper is the problem of task trees with deadlines~\cite{cohen2015estimating,CohenGW18}. Hierarchical planning is a well-established field in AI~\cite{thomas1988hierarchical,erol1994htn,erol1996complexity}, and is still relevant nowadays~\cite{alford2016hierarchical,xiao2017hierarchical}. A hierarchical plan is a method for representing problems of automated planning in which the dependency among tasks can be given in the form of networks, here we focus on hierarchical plans represented by task trees. The leaves in a task tree are \emph{primitive} actions (or tasks), and the internal nodes are either \emph{sequence} or \emph{parallel} actions. The plans we deal with are of stochastic nature, and the task duration is described as probability distributions in the leaf nodes. We assume that the distributions are independent but {\em not} necessarily identically distributed and that the random variables are discrete and have a finite support.

A sequence node denotes a series of tasks that should be performed consecutively, whereas a parallel node denotes a set of tasks that begin at the same time. A \emph{valid} plan is one that is fulfilled before some given \emph{deadline}, i.e., its \emph{makespan} is less than or equal to the deadline. The objective in this context is to compute the probability that a given plan is valid, or more formally computing $P(X<T)$, where $X$ is a random variable representing the makespan of the plan and $T$ is the deadline. The problem of finding the probability that a task tree satisfies a deadline is known to be NP-hard~\cite{mohring2001scheduling}. In fact, even the problem of summing a set of random variables is NP-hard~\cite{mohring2001scheduling}. This is an example of an explicitly given random variable that we need to estimate deadline meeting probabilities for.

@@

The rest of the paper is organised as follows. In Section~\ref{sec:relwork} we describe how our work relates to other algorithms and problems studied in the literature. In Section~\ref{sec:alg} we detail the proposed algorithm, analyse its properties, and prove the main theorem. In Section~\ref{sec:exp} we demonstrate how the proposed approach performs on the problem of estimating the probability of missing deadlines in series-parallel schedules on randomly generated random variables and compare it to alternative approximation approaches from the literature. The paper is concluded with a discussion and with ideas for future work in Section~\ref{sec:discussion}.

@@

In the technical level, the problem we study in this paper is similar to the problem of approximating a set of 2-D points by a step function. The study of this problem was motivated by query optimisation and histogram constructions in database management systems~\cite{applf12,applf13,applf14,applf17,applf18,Fournier2011} and computational geometry~\cite{diaz2001fitting,fournier2008fitting}. There are, however, two technically significant differences between the problem studied in the context of databases and the problem we analyse in this paper. The first difference is that in the context of approximation of random variables, the step function (which is the cumulative distribution function in our context)  must end with a value one, since we are dealing with random variables which sums to one. The second difference is that the first step is not counted because there is no need to put a value in the support of the approximated random variable to generate this first step. These cannot be addressed by adding a constant (two) to $m$ because the first step is always present and because the requirement to end with the value one, restricts the set of eligible step functions. See Section~\ref{sec:alg} for more details.

@@

Our study is also related to the work of Pavlikov and Uryasev~\cite{pavlikov2016cvar}, where a procedure for producing a random variable $X'$ that optimally approximates a random variable $X$ is presented. Their approximation scheme, achieved using linear programming, is designed for a different notion of distance called CVaR. The contribution of the present work in this context is that our method is direct, not using linear programming, thus allowing tighter analysis of time and memory complexities. Also, our method is designed for minimising the Kolmogorov distance that is more prevalent in applications. For comparison, in Section~\ref{sec:exp} we briefly discuss the performance of linear programming approach similar to the one proposed in~\cite{pavlikov2016cvar} for the Kolmogorov distance and compare it our algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An algorithm for optimal approximation}\label{sec:algEfficient}
In the scope of this section, let $X$ be a random variable such that $|X|=n$, and let  $0<m\leq n$. We present algorithms for finding optimal $m$-approximations of $X$. The first algorithm is $binsearchApprox(X,m)$ (Algorithm~\ref{alg:naive}) which runs in time of $O(n^2\log(n))$. The second algorithm, $linApprox(X,m)$ (Algorithm~\ref{alg:linear}), runs in time of $O(n\log(n))$. 
%We present both $binsearchApprox(X,m)$ and $linApprox(X,m)$ since Algorithm~\ref{alg:linear} is based on Algorithm~\ref{alg:naive} but carry major run-time improvement.

%We begin our dissection by presenting a useful fact that it is enough to limit our search to approximations $X'$s such that $\support(X') \subseteq \support(X)$:
%
%\begin{lemma}\label{lem:supContained}
%	For every random variable $X''$ there is a random variable $X'$ such that $\support(X') \subseteq \support(X)$ and $d_{K}(X,X')\leq d_{K}(X,X'')$.
%\end{lemma}
%
%\begin{proof}
%	Let $\{x_1 <\cdots<x_n\} = \support(X)$, and let $x_0 = -\infty, x_{n+1}=\infty$. Consider the random variable $X'$ whose probability mass function is
%	$f_{X'}(x_i) = P(x_{i-1} < X'' \leq x_i)$ for $i=1,\dots,n-1$,  $f_{X'}(x_n) = P(x_{n-1} < X'')$, and $F_{X'}(x)=0$ if $x\notin \support(X)$.  Since we only "shifted" the probability mass of $X''$ to the support of $X$, we have that $f_{X'}$ is a probability mass function and therefore $X'$ is well defined. 	
%	By construction, $|F_{X}(x_i)-F_{X'}(x_i)| = |F_{X}(x_i)-F_{X''}(x_i)|$ for every  $0 <  i < n$. For $i=n$ we have $|F_{X}(x_n)-F_{X'}(x_n)| = |1-1|=0$.
%	Since $|F_{X}(x)-F_{X'}(x)| = |F_{X}(x_i)-F_{X'}(x_i)|$ for every  $0\leq   i  \leq  n$  and $x_i<x<x_{i+1}$, we have that $d_K(X,X')=max_{i}|F_{X}(x_i)-F_{X'}(x_i)|\leq max_{i}|F_{X}(x_i)-F_{X''}(x_i)|\leq d_K(X,X'')$.
%\end{proof}

The algorithms presented in this section are using a dual problem in order to find the optimal $m$-approximation of $X$. The dual problem is: given a random variable $X$ and $0 \leq \varepsilon \leq 1$, find a new random variable $X'$ such that $d_k(X,X')\leq \varepsilon$ and $|X'|$ is minimal.

We do not assume anything about the representation of the random variable $X$ given as an input to the $dual(X,\varepsilon)$ algorithm. However, we transform the representation of every given $X$, to the form of sorted CDF as described in line 1 in the algorithm, which takes $O(nlog(n))$ run-time and has to done only once. The $dual(X,\varepsilon)$ gets as input a random variable $X$ and some $\varepsilon$, the random variable $X$ is represented as in line 1, $X_{CDF} =\{(x_i, c_i)\}_{i=1}^n$ such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$. In lines 3-4 we perform a single pass over $X_{CDF}$ starting from index 1 to find the last index where the cumulative distribution function of $X$ is less than $\varepsilon$, this index is now $f$. In lines 5-6 in the algorithm we start from the last index of $X_{CDF}$ which is $n$, and in decreasing order seek for the last index where the cumulative distribution function of $X$ is less than $1-\varepsilon$, this index is now $l$. The last part of the algorithm is a pass on all elements in $X_{CDF}$ between the indices $f$ and $l$ in order to construct the $m$-approximation random variable of $X$, $X'$. In lines 9-10 we check if the difference between the first element in the current set of indices (which we call $b$) and the last element in the current set of indices (which we call $e+1$) is less that $2\varepsilon$. If yes, we add the pair $\{(x_{b}, (c_b+c_e)/2 - s) \}$ to the set $S$. Otherwise, we add one to the counter of $e$. Note, the reason we need $s$ is to return the random variable $X'$ in a PDF representation since our main motivation is sum of random variables, this part takes $O(n)$ run-time.

\begin{algorithm}
	\DontPrintSemicolon
	%Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$.\;
    $f \gets 0$, $l \gets n$+1\; 
    \While{$c_{f +1} \leq \varepsilon$}{$f \gets f + 1$}

    \While{$c_{l -1} \geq  1-\varepsilon$}{$l \gets l -1$}

@@

\begin{theorem}\label{the:complexityDual}
	The $dual(X,\varepsilon)$ algorithm runs in time $O(n)$, using $O(n)$ memory where $n=|X|$.
\end{theorem}
\begin{proof}
	This algorithm describes a single pass over the random variable $X$. Lines 3-4 and lines 5-6 are easy to follow, each takes $O(n)$ in the worst case. Lines 8-12 also describe one pass over $X$ since the counter $e$ is updated to $e+1$ at most $n$ times. All together we get run-time complexity of $O(3n)$ = $O(n)$. We are constructing the set $S$ which is of size $n$ in the worst case, therefore, memory complexity is also $O(n)$. 
\end{proof}

The first technique we present is the $binsearchApprox(X,m)$ algorithm which is based on the work done by D{\'i}az-B{\'a}nez and Mesa~\cite{diaz2001fitting}. There are couple of significant changes between our $binsearchApprox(X,m)$ algorithm and the algorithm suggested by D{\'i}az-B{\'a}nez and Mesa~\cite{diaz2001fitting} addressing the differences presented in the Related Work section~\ref{sec:relwork}. The algorithm $binsearchApprox(X,m)$ gets as input a random variable $X$ and some number $m$. Again, we do not mind the original representation of $X$ since we can transform it to a sorted CDF representation in $O(nlog(n))$ run-time as in line 1 which we call $X_{CDF}$. In lines 3-8 we compute all possible errors, in other words, all possible $d_{K}(X,X')$ such that $X'$ is an approximation of $X$ and $\support(X') \subseteq \support(X)$. The error is just the difference between the CDF values of every two elements in $X_{CDF}$. After computing the set $E$, we sort it (line 9) in order to perform a binary search in the last part of the algorithm. In lines 10-19 we perform a binary search over the set $E$, in every step of the binary search we run the $X' = dual(X,\varepsilon)$ algorithm. If the size of $|X'|>m$ then we know that the error $\varepsilon$ is too small and need to extend the search in the right side of $E$, if the size of $|X'|<m$ then we know that the error $\varepsilon$ is too big and need to extend the search in the left side of $E$, otherwise we found the correct $m$. We run the search twice to handle the extreme case where $m \notin \{\forall \varepsilon\in E: |dual(X,\varepsilon)|\}$ and return the smallest $m'$ closest to $m$.

@@

    \For{$i \gets 1 \textbf{ to } 2$}{
        $l\gets 1$, $r \gets n'$, $k \gets 1$ \;
        \Do{$k \neq k'$}{
                $k' \gets k$,$k \gets \lceil (l + r)/2 \rceil$\;
        		$X' \gets dual(X,e_k)$, $m' \gets |X'|$\;
        		\lIf{$m' < m$}{
        		    $l \gets k$
        		}
                \lIf{$m' > m$}{
        		    $r \gets k-1$
        		}
                \lIf{$m' = m$}{
        		    $r \gets k$
        		}
        }
        $m \gets m'$\;
    }
    \Return{X'}

	\caption{$binsearchApprox(X,m)$}   
	\label{alg:naive}
\end{algorithm}

\begin{theorem}\label{the:complexityDual}
	The $binsearchApprox(X,m)$ algorithm runs in time $O(n^2log(n))$, using $O(n^2)$ memory where $n=|X|$.
\end{theorem}
\begin{proof}
	In the first part of the algorithm, lines 2-8, we construct the set $E$ which takes $O(n^2)$ run-time. In the second part of the algorithm, line 9, we sort the set $E$ which takes $O(n^2log(n^2)) = O(n^2log(n))$  run-time. The third part of the algorithm, lines 10-19, describes a binary sort over the set $E$ where in each step of the sorting we run the $dual$ algorithm, which takes $O(n log(n^2)) = O(n log(n))$ run-time. We run this part twice then we get $O(2n log(n))$. All together, the run-time complexity is $O(n^2+n^2log(n)+2n log(n)) = O(n^2log(n))$ and the memory complexity is for storing the set $E$ which is $O(n^2)$.
\end{proof}

As a step towards $linApprox(X,m)$ algorithm, we present the algorithm $saddlebackApprox(X,m)$ that takes $O(n^2)$ run-time, this technique is very similar to $linApprox(X,m)$ algorithm. The algorithm $saddlebackApprox(X,m)$ gets as input a random variable $X$ and some number $m$. Again, we represent $X$ as a sorted CDF in $O(nlog(n))$ run-time as in line 1 and call it $X_{CDF}$. In line 4 We describe implicitly the sorted matrix $E$, and in lines 3-7 we perform a search over this matrix, in every step of the search we run the $X' = dual(X,\varepsilon)$ algorithm. If the size of $|X'|>m$ then we know that the error $e$ is too small and need to extend the search in the lower side of the matrix, if the size of $|X'|<m$ then we know that the error $e$ is too big and need to extend the search in the right side of the matrix and store this $e$ in $S$. We store in set $S$ all possible candidates for the correct $e$ and in line 8 we get the maximal $m$ with the minimal $e$ value.

\begin{algorithm}
	\DontPrintSemicolon
	Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$.\;
    $i \gets 1$, $j \gets \lceil n/2 \rceil$, $S \gets \emptyset$\;

	\While{$i < \lfloor n/2 \rfloor + 1 \wedge j \geq 2$}
	{
% 		$ e \gets \begin{cases}
% 			c_j   & \text{if } i=\lfloor n/2 \rfloor+1; \\
% 			1-c_i & \text{if } j=\lceil n/2 \rceil; \\
% 			(c_{n+1-j} {-} c_i)/2  & \text{otherwise}
% 		\end{cases}$\;

		$X' \gets dual(X,e_{i,j})$, $m' \gets |X'|$\;

@@

	\caption{$saddlebackApprox(X,m)$}   
	\label{alg:saddleback}
\end{algorithm}

@@

		$w \gets 2^{\lceil \log_2(n/2) \rceil}$,
		$h \gets 2^{\lceil \log_2(n/2+1) \rceil}$ \;

		$S \gets  \{ ((1, 1),(w,h)) \}$, $S' \gets \emptyset$\;

@@

			\For{$ e \in \{ e^-,e^+ \}$} {
				$X' \gets dual(X,e)$, $m' \gets |X'|$\;

@@

		\caption{$linApprox(X,m)$}   
		\label{alg:linear}
	\end{algorithm}

	\begin{procedure}
		\DontPrintSemicolon
		$j^- \gets \lfloor (j_1+j_2)/2 \rfloor$,
		$j^+ \gets \lceil (j_1+j_2)/2 \rceil$\;
		$i^- \gets \lfloor (i_1+i_2)/2 \rfloor$,
		$i^+ \gets \lceil (i_1+i_2)/2 \rceil$\;

@@

		\caption{{$split(((i_1,j_1),(i_2,j_2)))$}()}
		\label{proc:split}
	\end{procedure}

>>>>

\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true,factor=1100,stretch=10,shrink=100]{microtype}

@@

\DeclareMathOperator{\supp}{support}
\DeclareMathOperator{\support}{support}
\DeclareMathOperator{\Trim}{Trim}
\DeclareMathOperator{\LTrim}{LTrim}
\SetKwFunction{mEqOne}{mEqOne} 
\DeclareMathOperator{\KlmApprox}{linApprox}
\DeclareMathOperator{\BinsApprox}{binsApprox}
\DeclareMathOperator{\OptTrim}{OptTrim}

  \pdfinfo{
	/Title (Optimal Approximation of Discrete Random Variables for Estimation of Probabilities of Missing Deadlines
	)
	/Author ()}
\setcounter{secnumdepth}{0}

@@

Various approaches for approximation of probability distributions are studied in the literature~\cite{PS77,AMCR83,vidyasagar2012metric,cohen2015estimating,pavlikov2016cvar,CohenGW18}. 
These approaches vary in the types random variables considered, how they are represented, and in the criteria used for evaluation of the quality of the approximations. In this paper we propose an approach for compressing the probability mass function of a random variable $X$ such that the errors added to queries such as $Pr(X\leq t)$, for  any $t>0$, is minimal. In other words, we minimise the Kolmogorov distance between the approximation and the original random variable. Our motivation for this is estimation of the probability for missing deadlines, as described, e.g., in Cohen et al.~\cite{cohen2015estimating,CohenGW18} and in~\cite{Kashef18}. Specifically, when $X$ represents the probability distribution of the time to complete some complex schedule and we cannot afford to maintain the full table of its probability mass function, we propose an algorithm for producing a smaller table, whose size can be specified, such that probabilities for missing deadlines are preserved as much as possible.

@@

The case study examined in this paper is the problem of task trees with deadlines~\cite{cohen2015estimating,CohenGW18}. A hierarchical plan is a method for representing problems of automated planning in which the dependency among tasks can be given in the form of networks, here we focus on hierarchical plans represented by task trees. The leaves in a task tree are \emph{primitive} actions (or tasks), and the internal nodes are either \emph{sequence} or \emph{parallel} actions. The plans we deal with are of stochastic nature, and the task duration is described as probability distributions in the leaf nodes. We assume that the distributions are independent but {\em not} necessarily identically distributed and that the random variables are discrete and have a finite support.

A sequence node denotes a series of tasks that should be performed consecutively, whereas a parallel node denotes a set of tasks that begin at the same time. A \emph{valid} plan is one that is fulfilled before some given \emph{deadline}, i.e., its \emph{makespan} is less than or equal to the deadline. The objective in this context is to compute the probability that a given plan is valid, or more formally computing $P(X<T)$, where $X$ is a random variable representing the makespan of the plan and $T$ is the deadline. The problem of finding the probability that a task tree satisfies a deadline is known to be NP-hard~\cite{cohen2015estimating}. In fact, even the problem of summing a set of random variables is NP-hard~\cite{mohring2001scheduling}. This is an example of an explicitly given random variable that we need to estimate deadline missing probabilities for.

@@

The rest of the paper is organised as follows. In Section~\ref{sec:relwork} we describe how our work relates to other algorithms and problems studied in the literature. In Section~\ref{sec:algEfficient} we detail the proposed algorithm, analyse its properties, and prove the main theorems. In Section~\ref{sec:exp} we demonstrate how the proposed approach performs on the problem of estimating the probability of missing deadlines in series-parallel schedules on randomly generated random variables and compare it to alternative approximation approaches from the literature. The paper is concluded with a discussion and with ideas for future work in Section~\ref{sec:discussion}.

@@

In the technical level, the problem we study in this paper is similar to the problem of approximating a set of 2-D points by a step function. The study of this problem was motivated by query optimisation and histogram constructions in database management systems~\cite{applf12,applf13,applf14,applf17,applf18,Fournier2011} and computational geometry~\cite{diaz2001fitting,fournier2008fitting}. There are, however, two technically significant differences between the problem studied in the context of databases and the problem we analyse in this paper. The first difference is that in the context of approximation of random variables, the step function (which is the cumulative distribution function in our context)  must end with a value one, since we are dealing with random variables which sums to one. The second difference is that the first step is not counted because there is no need to put a value in the support of the approximated random variable to generate this first step. These cannot be addressed by adding a constant (two) to $m$ because the first step is always present and because the requirement to end with the value one, restricts the set of eligible step functions. See Section~\ref{sec:algEfficient} for more details.

@@

Our study is also related to the work of Pavlikov and Uryasev~\shortcite{pavlikov2016cvar}, where a procedure for producing a random variable $X'$ that optimally approximates a random variable $X$ is presented. Their approximation scheme, achieved using linear programming, is designed for a different notion of distance called CVaR. The contribution of the present work in this context is that our method is direct, not using linear programming, thus allowing tighter analysis of time and memory complexities. Also, our method is designed for minimising the Kolmogorov distance that is more prevalent in applications. For comparison, in Section~\ref{sec:exp} we briefly discuss the performance of linear programming approach similar to the one proposed in~\cite{pavlikov2016cvar} for the Kolmogorov distance and compare it our algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algorithms for optimal approximation}\label{sec:algEfficient}
%In the scope of this section, let $X$ be a random variable ket $m$ be such that  $0<m\leq |X|$. We present algorithms for finding optimal $m$-approximations of $X$.

We begin with presenting an algorithm for solving a problem that is dual to the $m$-approximation problem.  The dual problem is: given a random variable $X$ and $0 \leq \varepsilon \leq 1$, find a new random variable $X'$ such that $d_k(X,X')\leq \varepsilon$ and $|X'|$ is minimal.

We assume that the input to Algorithm~\ref{alg:dual} that solves the dual problem is a representation of the variable $X$ as a sorted CDF, i.e.,  as an array $X_{CDF}=\{(x_i, c_i)\}_{i=1}^n$ such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$. In lines 2 we perform a single pass over $X_{CDF}$ starting from index 1 to find the last index where the cumulative distribution function of $X$ is less than $\varepsilon$, this index is now $f$. In line 3 we start from the last index of $X_{CDF}$ which is $n$, and in decreasing order seek for the last index where the cumulative distribution function of $X$ is less than $1-\varepsilon$, this index is now $l$. The last part of the algorithm is a pass on all elements in $X_{CDF}$ between the indices $f$ and $l$ in order to construct the $m$-approximation random variable of $X$, $X'$. In lines 6-7 we check if the difference between the first element in the current set of indices (which we call $b$) and the last element in the current set of indices (which we call $e+1$) is less than $2\varepsilon$. If yes, we add the pair $\{(x_{b}, (c_b+c_e)/2 - s) \}$ to the set $S$. Otherwise, we add one to the counter of $e$.

\begin{algorithm}
	\DontPrintSemicolon
	%Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$.\;
    $f \gets 0$, $l \gets n$+1\; 
    \lWhile{$c_{f +1} \leq \varepsilon$}{$f \gets f + 1$}

    \lWhile{$c_{l -1} \geq  1-\varepsilon$}{$l \gets l -1$}

@@

\begin{theorem}\label{the:correctnessDual}
	If  $\{(x_i, c_i)\}_{i=1}^n$ is such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$ then  
    $$dual(\{(x_i, c_i)\}_{i=1}^n,\varepsilon) \in \underset{X'\in \bar{\mathcal{B}}(X,\varepsilon)}{\operatorname{arg\,min}}\, |X'|$$  where $\bar{\mathcal{B}}(X,\varepsilon)=\{X'\colon d_K(X,X')\leq \varepsilon\}$.
\end{theorem}
\begin{proof}
The number of level sets of the CDF of $X'$ is minimal: (1) The first and the last sets are of maximal length; (2) By construction, an extension of any of the other sets to the right will generate a random variable whose Kolmogorov distance from $X$ is bigger than $\varepsilon$; and (3) Extension of a level set to the left may either leave the number of level sets unchanged or combine it with the previous set, which will enlarge the Kolmogorov distance because it is equivalent to extending the previous set to the right.
\end{proof}

\begin{theorem}\label{the:complexityDual}
	$dual(\{(x_i, c_i)\}_{i=1}^n,\varepsilon)$ runs in time $O(n)$, using $O(n)$ memory.
\end{theorem}
\begin{proof}
This algorithm describes a single pass over $\{(x_i, c_i)\}_{i=1}^n$. Line 2 and line 3 are easy to follow, each takes $O(n)$ in the worst case. Lines 5-9 also describe a single pass since the counter $e$ is updated to $e+1$ at most $n$ times. All together we get run-time complexity of $O(n)$. We are constructing the set $S$ which is of size $n$ in the worst case, therefore, memory complexity is also $O(n)$. 
\end{proof}

The first technique we present is the $binsApprox(X,m)$ algorithm which is based on the work done by D{\'i}az-B{\'a}nez and Mesa~\cite{diaz2001fitting}. There are couple of significant changes between our $binsApprox(X,m)$ algorithm and the algorithm suggested by D{\'i}az-B{\'a}nez and Mesa~\shortcite{diaz2001fitting} addressing the differences presented in the Related work section. The algorithm $binsApprox(X,m)$ gets as input a random variable $X$ and some number $m$. Again, we do not mind the original representation of $X$ since we can transform it to a sorted CDF representation in $O(nlog(n))$ run-time as in line 1 which we call $X_{CDF}$. In lines 3-8 we compute all possible errors, in other words, all possible $d_{K}(X,X')$ such that $X'$ is an approximation of $X$ and $\support(X') \subseteq \support(X)$. The error is just the difference between the CDF values of every two elements in $X_{CDF}$. After computing the set $E$, we sort it (line 9) in order to perform a binary search. In lines 10-19 we perform a binary search over the set $E$, in every step of the binary search we run the $X' = dual(X,\varepsilon)$ algorithm. If the size of $|X'|>m$ then we know that the error $\varepsilon$ is too small and need to extend the search in the right side of $E$, if the size of $|X'|<m$ then we know that the error $\varepsilon$ is too big and need to extend the search in the left side of $E$, otherwise we found the correct $m$. We run the search twice to handle the extreme case where $m \notin \{\forall \varepsilon\in E: |dual(X,\varepsilon)|\}$ and return the smallest $m'$ closest to $m$.

@@

    \For{$i \gets 1 \textbf{ to } 2$}{
        $l\gets 1$, $r \gets n'$, $k \gets 1$, $k' \gets 0$ \;
        \While{$k \neq k'$}{
                $k' \gets k$,$k \gets \lceil (l + r)/2 \rceil$\;
        		$m' \gets |dual(X,e_k)|$\;
        		\lIf{$m' < m$}{
        		    $l \gets k$
        		}
                \lIf{$m' > m$}{
        		    $r \gets k-1$
        		}
                \lIf{$m' = m$}{
        		    $r \gets k$
        		}
        }
        $m \gets m'$\;
    }
    \Return{$dual(X,e_k)$}

	\caption{$binsApprox(X,m)$}   
	\label{alg:naive}
\end{algorithm}

\begin{theorem}\label{the:correctnessBinsearch}

    $binsApprox(X,m) \in \underset{X' \in \bar{\mathcal{B}}(X,m) }{\operatorname{arg\,min}}\, d_K(X,X')$ where $\bar{\mathcal{B}}(X,m)=\{X'\colon |X'| \leq m\}$.
\end{theorem}
\begin{proof}
Since lines 10-18 is a binary search of the smallest $e_k\in E$ such that $|dual(X,e_k)|\leq m$, we only need to prove that $E=E'=\{ d_K(X,dual(X,m)) \colon m=1,\dots,n\}$. To see that this is true, note that every element in $E'$ corresponds to a distance of a level set of the CDF of $dual(X,m)$ from the CDF of $X$. Line 6 of the algorithm adds all the distances from level sets of height zero, Line 8 adds the distances from level sets of heigt one, and Line 7 adds all the distances from all other possible level sets.
\end{proof}

\begin{theorem}\label{the:complexityBinsearch}
	The $binsApprox(X,m)$ algorithm runs in time $O(n^2log(n))$, using $O(n^2)$ memory where $n=|X|$.
\end{theorem}
\begin{proof}
	In the first part of the algorithm, lines 2-8, we construct the set $E$ which takes $O(n^2)$ run-time. In the second part of the algorithm, line 9, we sort the set $E$ which takes $O(n^2log(n^2)) = O(n^2log(n))$  run-time. The third part of the algorithm, lines 10-19, describes a binary sort over the set $E$ where in each step of the sorting we run the $dual$ algorithm, which takes $O(n log(n^2)) = O(n log(n))$ run-time. We run this part twice then we get $O(2n log(n))$. All together, the run-time complexity is $O(n^2+n^2log(n)+2n log(n)) = O(n^2log(n))$ and the memory complexity is for storing the set $E$ which is $O(n^2)$.
\end{proof}

Towards an improved algorithm let us introduce the matrix $E=(e_{i,j})_{i,j=1}^\infty$ defined by:
	$$e_{i,j} = \begin{cases}
	1-c_{n+1-j}   & \text{if } j \leq n \wedge  i = n ; \\
	c_i & \text{if } i \leq n \wedge  j = n + 1; \\
	( c_i - c_{n+1-j})/2  & \text{if } i < n \wedge j \leq n \wedge i{+}j\geq n; \\
	0  & \text{if }  i{+}j < n; \\
	1  & \text{otherwise}.
	\end{cases}$$
Where $c_1,\dots,c_n$ are as in the first line of Algorithm~\ref{alg:naive}. It is easy to see that the set of values in this matrix are the elements of the set $E$ in Algorithm~\ref{alg:naive}. An additional useful fact is that $E$ is a sorted matrix:
\begin{lemma}\label{the:sortedMatrix}
	If $i\leq i'$ and $j\leq j'$ then $e_{i,j}\leq e_{i',j'}$.
\end{lemma}
\begin{proof}
CDF representation of a random variable is a monotonically increasing function. Therefore, $\forall i,j$, $ 0\leq c_i - c_{n+1-j}\leq c_i - c_{n+1-(j+1)} $ and $ \forall i,j$, $0 \leq c_i - c_{n+1-j}\leq c_{i+1} - c_{n+1-j} $, moreover, 0 is the minimal value of $E$. The elements in the last row and the last column in $E$ are keeping the terms of sorted matrix. It suffices to compare row $n$ with row $n-1$ and show that $1-c_{n+1-j}\geq ( c_{n-1} - c_{n+1-j})/2$. After some manipulation we get that $2-c_{n+1-j}\geq c_{n-1}$ which is true because $0 \leq c\leq 1$. Next, it is suffice to compare column $n$ with only column $n+1$ and show that $c_i\geq ( c_{i} - c_{n+1-n})/2$. After some manipulation we get that $2c_i\geq c_{i}-c_1$ which is true because $0 \leq c_1 \leq c_i\leq 1$.
\end{proof}

The fact that this matrix is sorted allows us to use the saddleback search algorithm listed as Algorithm~\ref{alg:saddleback}. The algorithm starts at the top right entry of the matrix $(e_{i,j})_{i=1..n,j=1..(n+1)}$ and traverses it as follows. If it hits an entry $e$ such that $|dual(X,e)| \leq m$ it goes left, otherwise it goes down. This assures that the minimal $e$ such that $|dual(X,e)| \leq m$ is visited after at most $n+1$ steps. Data is collected along the traversal and the optimal random variable is found by brute-force.

\begin{algorithm}
	\DontPrintSemicolon
	Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$.\;
    $i \gets 1$, $j \gets n$+1, $S \gets \emptyset$\;

	\While{$i < n \wedge j \geq 1$}
	{	
		$m' \gets |dual(X,e_{i,j})|$\;

@@

	\caption{$sdlbkApprox(X,m)$}   
	\label{alg:saddleback}
\end{algorithm}

\begin{theorem}\label{the:correctnessSaddleback}

    $sdlbkApprox(X,m) {\in} \underset{X' \in \bar{\mathcal{B}}(X,m) }{\operatorname{arg\,min}}\, d_K(X,X')$ where $\bar{\mathcal{B}}(X,m)=\{X'\colon |X'| \leq m\}$.
\end{theorem}
\begin{proof}
The algorithm traverses all the frontier between those $e$s that satisfy $|dual(X,e)| \leq m$ and those that do not satisfy it. Since all the entries that satisfy the condition are recorded in $S$ and considered in the brute-force phase in line 7, the minimal satisfying $e$ is found.
\end{proof}

\begin{theorem}\label{the:complexitySaddleback}
	The $sdlbkApprox(X,m)$ algorithm runs in time $O(n^2)$, using $O(n)$ memory where $n=|X|$.
\end{theorem}
\begin{proof}
	At each step of the loop in lines 3-6, either $j$ decreases or $i$ increases, thus the loop can be executed at most $2n$ times. Since we execute $dual$ once ion a loop round, the total time complexity is $O(n^2)$. Storing visited states in $S$ on line 6 requires  $O(n)$ memory.
\end{proof}

The problem with the saddleback algorithm is that it needs to run $dual$ at every step so it has a quadratic time complexity. Since we cannot find the required entry of the matrix in less than $n$ steps, we can only reduce the complexity by proposing an algorithm that does not executes $dual$ in all the steps, only in $\log(n)$ of them. Such an algorithm, based on Section~2.1 in~\cite{Fournier2011}, is listed as Algorithm~\ref{alg:linear}. The algorithm maintains a set $S$ of sub-matrices. At each round of its execution, each sub-matrix is split to four and then about three quarters of the matrix are discarded. At the end, at most four scalar matrices remain containing the index of the entry we seek.

@@

		$S \gets  \{ ((1, 1),(n,n+1)) \}$, $S' \gets \emptyset$\;

@@

			\For{$ e \in \{ e^-,e^+ \}$} {
				$m' \gets |dual(X,e)|$\;

@@

		\caption{$linApprox(X,m)$}   
		\label{alg:linear}
	\end{algorithm}
{
	\begin{algorithm}
		\DontPrintSemicolon

		$j^- \gets \lfloor (j_1+j_2)/2 \rfloor$,
		$j^+ \gets \lceil (j_1+j_2)/2 \rceil$\;
		$i^- \gets \lfloor (i_1+i_2)/2 \rfloor$,
		$i^+ \gets \lceil (i_1+i_2)/2 \rceil$\;

@@

		\caption{$split(((i_1,j_1),(i_2,j_2)))$}
		\label{proc:split}
	\end{algorithm}
}

\begin{theorem}\label{the:correctnessLinear}
